---
output: 
  github_document:
    toc: false
    fig_width: 7
    fig_height: 5
---
  
  <!-- README.md is generated from README.Rmd. Please edit that file -->
  
```{r, echo = FALSE, warning=FALSE, message=FALSE}
options(digits = 2)

knitr::opts_chunk$set(
  collapse = TRUE,
  fig.path = "man/figures/",
  fig.width = 7,
  fig.height = 5,
  comment = "#>",
  dev = "png",
  dpi = 300
)
```
  
# {scatter}
  
This package provides functions for quickly calculating measures of model prediction accuracy and creating enhanced scatterplots. The scatterplots can include text summarizing the agreement metrics (e.g., R², RMSE, MAPE) between two plotted variables, with support for grouped data and faceting. Agreement metrics are sourced from the [{yardstick}](https://yardstick.tidymodels.org/index.html) package and can be further customized by the user.

# Why?

Scatterplots are one of the most frequently used visualizations in my daily work. While creating a scatterplot is straightforward, I wanted a version that would meet these specific needs:

- Ensure the plot is always square.
- Include a 1:1 reference line.
- Provide an optional text panel showing agreement statistics, allowing an easy customization of the metrics included.
- Seamlessly handle grouped data, creating faceted plots with minimal effort.


# Installation

```{r, eval=FALSE}
devtools::install_github("ptompalski/scatter")
library(scatter)
```


# Examples

Simple scatterplot with R² and RMSE

```{r, warning=FALSE, message=FALSE, fig.height=3.5}
library(dplyr)
library(scatter)

# some fake data
df <- 
tibble(
  truth = c(rnorm(150, 10, 2)),
  estimate = truth + rnorm(150,0, 1),
  group = rep(c("A", "B", "C"), each = 50),          
  group2 = rep(c("D1", "D2"), each = 75),          
)

scatter(df, truth, estimate, metrics=list(rsq,rmse))
```



Scatterplot for grouped data with agreement metrics
```{r, warning=FALSE, message=FALSE, fig.height=3.5}
df %>%
  group_by(group) %>%
  scatter(truth, estimate, metrics=list(rsq,rmse,mape,msd))
```



Scatterplot for grouped data with agreement metrics positioned outside the plots.
```{r, warning=FALSE, message=FALSE, fig.height=3.5}
df %>%
  group_by(group) %>%
  scatter(truth, estimate, metrics=list(rsq,rmse), metrics_position = "outside")
```


## Axis orientation: observed vs. predicted

By default, `scatter()` places observed (truth) values on the x-axis and predicted (estimate) values on the y-axis. This follows recommendations in the modelling literature, but users can easily switch back to the alternative convention (predicted on x) using `swap_axes = TRUE`.

The most influential position on this topic, published by [Piñeiro et al. (2008)](https://www.sciencedirect.com/science/article/pii/S0304380008002305), argues that observed values should be placed on the x-axis to improve the interpretability of regression diagnostics and model evaluation, based on the idea that predictions are conditioned on observations. However, [Pauwels et al. (2019)](https://www.sciencedirect.com/science/article/pii/S0304380019303102?via%3Dihub) subsequently revisited this argument and pointed out that parts of this reasoning are conceptually flawed, showing that the choice of axis orientation does not have a universally correct solution and can influence interpretation depending on the modelling context and analytical objective.

Importantly, axis orientation in `scatter()` affects only the visualization; agreement metrics are always computed using the original (truth, estimate) pairing.